---
title: PlotCat
---

PlotCat turns a plotting chunk into an exercise. Students see the finished plot, write R or Python in the browser, and compare their SVG with the target. Their code and results stay in the browser.

## Install

Copy `_extensions/plotcat` into your Quarto project and add the filter to `_quarto.yml`:

```yaml
filters:
  - plotcat
```

R exercises use WebR. Python exercises use Pyodide. Both runtimes load after a student clicks **Run**.

## Write an exercise

Wrap one target chunk in a `.plotcat` div. Give the div a unique `id`.

````markdown
::: {.plotcat id="cars"}
```{r}
plot(cars)
```
:::
````

Add a second chunk for starter code. Mark it `eval: false` so Quarto does not run it while rendering the page.

````markdown
::: {.plotcat id="cars-with-starter"}
```{r}
plot(cars, main = "Stopping distance")
```

```{r}
#| eval: false
plot(cars)
```
:::
````

The first chunk produces the target SVG during `quarto render`. PlotCat removes its source from the HTML. The optional second chunk appears in the editor but does not run during rendering.
