# PlotCat

`plotcat` is a Quarto extension for plot recreation exercises. Authors provide a target plot and optional starter code. Students write R or Python in the browser, render a plot with WebR or Pyodide, and compare it with the target.

See the [examples](examples.qmd) for exercises using tinyplot, ggplot2, Matplotlib, and plotnine.

## Installation

Add the extension to a Quarto project:

```bash
quarto add VisruthSK/PlotCat
```

Then enable the filter in a document or project config:

```yaml
filters:
  - plotcat
```

## Example

Write the target plot as the first executable chunk inside a `.plotcat` Div. The Div needs a unique `id`.

````markdown
::: {.plotcat id="cars-tinyplot"}
```{r}
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```
:::
````

Add a second chunk to give students starter code. Set `eval: false` so Quarto does not run the starter while rendering the document.

````markdown
::: {.plotcat id="cars-tinyplot-starter"}
```{r}
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```

```{r}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars
)
```
:::
````

The first chunk runs during `quarto render` and becomes the target SVG. Its source code is omitted from the rendered HTML. The second chunk appears in the editor.

PlotCat accepts R and Python chunks. Both chunks in an exercise must use the same language. Only `id` is supported on the `.plotcat` Div.

## Development

Run the unit tests and render the two-page example site:

```bash
pnpm test
quarto render
```

PlotCat was inspired by [ggplot2 Battles](https://github.com/MikeLydeamore/ggplot2-battles).
