---
title: Reference
format:
  html:
    toc: true
    toc-depth: 3
---

Configuration options, chunk markup syntax, and CSS theme variables for the `plotcat` extension.

## Document Setup

Enable the filter in your document header or project `_quarto.yml`:

```yaml
filters:
  - plotcat
```

PlotCat requires interactive HTML output. It runs R code via WebR and Python code via Pyodide.

## Exercise Markup

Wrap plot recreation exercises in a `::: {.plotcat}` container div. Each exercise requires a target code chunk and an optional starter code chunk.

````markdown
::: {.plotcat}
```{r}
#| eval: false
ggplot2::ggplot(cars, ggplot2::aes(speed, dist)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = "Stopping distance")
```

```{r}
#| eval: false
ggplot2::ggplot(cars, ggplot2::aes(speed, dist)) +
  ggplot2::geom_point()
```
:::
````

### Markup Rules

- Set `#| eval: false` on both chunks. PlotCat executes R and Python code in the browser rather than during document rendering.
- The first chunk defines the target plot. PlotCat renders it in WebR or Pyodide and encrypts the target code in the HTML.
- The second chunk populates the student code editor. If omitted, the editor starts blank.
- Both chunks in an exercise must use the same language (`r`/`webr` or `python`/`pyodide`).
- Assign custom exercise IDs on the container div: `::: {.plotcat #scatter-exercise}`. If omitted, PlotCat auto-generates IDs (`exercise-1`, `exercise-2`).

## CSS Variables

PlotCat exposes CSS custom properties on `.plotcat` containers. Override these variables globally in your stylesheet or locally on specific containers.

| Variable | Default Value | Description |
| :--- | :--- | :--- |
| `--plotcat-border` | `color-mix(in srgb, var(--bs-body-color, #212529) 20%, var(--bs-body-bg, #ffffff))` | Outer border color for the exercise card and internal divider lines. |
| `--plotcat-accent` | `var(--bs-primary, #0d6efd)` | Primary brand color used for action buttons, focus highlights, score badges, active mode toggles, and the wipe slider handle. |
| `--plotcat-accent-rgb` | `var(--bs-primary-rgb, 13, 110, 253)` | Space-separated RGB values of `--plotcat-accent`, used for focus ring and badge background opacity channels. |
| `--plotcat-bg` | `var(--bs-body-bg, #ffffff)` | Card background color and target/student plot container background. |
| `--plotcat-nested-bg` | `color-mix(in srgb, var(--bs-body-color, #212529) 6%, var(--bs-body-bg, #ffffff))` | Surface color for the code editor textarea and the comparison control bar. |
| `--plotcat-shadow-color` | `rgb(var(--bs-emphasis-color-rgb, 0, 0, 0))` | Base shadow color used for card drop shadows and handle depth. |
| `--plotcat-wipe` | `50%` | Position percentage of the wipe comparison slider handle (managed dynamically by user dragging). |
| `--plotcat-score-hue` | `120` | Dynamic HSL hue angle (0 to 120) set automatically after plot comparison to color the score pill. |

### Bootstrap Variable Integration

PlotCat automatically inherits global colors from Quarto Bootstrap themes:

- `--bs-body-color`: Primary body text color applied to text, labels, and editor code.
- `--bs-body-bg`: Background color used as the baseline for card and plot surfaces.
- `--bs-primary`: Main theme brand color mapped to `--plotcat-accent`.
- `--bs-primary-rgb`: RGB tuple mapped to `--plotcat-accent-rgb`.
- `--bs-font-monospace`: Font stack used inside the code editor `<textarea>`.
- `--bs-body-font-family`: Font stack used for headers, labels, and buttons.
- `--bs-secondary-color`: Text color for muted comparison labels and status messages.
- `--bs-danger`: Text color applied when code execution returns an error.

## Custom Styling Examples

### Theme Accent Override

Change PlotCat accent color globally in a custom CSS file:

```css
.plotcat {
  --plotcat-accent: #059669;
  --plotcat-accent-rgb: 5, 150, 105;
}
```

### Scoped Container Styling

Adjust code editor height or card layout for specific exercise blocks:

```css
#scatter-exercise .plotcat__textarea {
  height: 16rem;
}

.plotcat.custom-card {
  --plotcat-bg: #f8fafc;
  --plotcat-border: #cbd5e1;
}
```
