---
title: Styling
format:
  html:
    toc: true
---

PlotCat uses modern CSS variables for styling. You can customize the look of all exercises globally, or style individual exercises by applying a custom class.

## CSS Variables

PlotCat defines these variables on the `.plotcat` class. Override them in your custom stylesheet or in a `<style>` block:

| Variable | Description | Default |
|---|---|---|
| `--plotcat-accent` | The highlight color for running status, score badges, and selected controls | `var(--bs-primary, #0d6efd)` |
| `--plotcat-border` | The border color for the panel and inputs | `var(--bs-border-color, #dee2e6)` |
| `--plotcat-bg` | The background color of the main container and plot wrappers | `var(--bs-body-bg, #ffffff)` |
| `--plotcat-nested-bg` | The background color of code textareas and compare control bar | `var(--bs-tertiary-bg, #f8f9fa)` |

## Example: Fancy Card Style

You can define a custom class to make exercises stand out. For example, add this to a custom stylesheet or a `<style>` block:

```css
.fancy-plotcat {
  --plotcat-accent: #8e44ad; /* Purple accent */
  --plotcat-border: #d2b4de;
  background: linear-gradient(135deg, #f5eef8, #ebdef0);
  border-radius: 1rem;
  box-shadow: 0 8px 30px rgba(142, 68, 173, 0.08);
}
.fancy-plotcat .plotcat__plot {
  border-radius: 0.75rem;
  border-color: #d2b4de;
  box-shadow: 0 4px 10px rgba(0, 0, 0, 0.04);
}
.fancy-plotcat .plotcat__button {
  border-radius: 2rem;
}
.fancy-plotcat .plotcat__compare-options {
  border-color: #d2b4de;
  background: #f5eef8;
}
```

```{#fancy-style-css =html}
<style>
.fancy-plotcat {
  --plotcat-accent: #8e44ad;
  --plotcat-border: #d2b4de;
  background: linear-gradient(135deg, #f5eef8, #ebdef0);
  border-radius: 1rem;
  box-shadow: 0 8px 30px rgba(142, 68, 173, 0.08);
}
.fancy-plotcat .plotcat__plot {
  border-radius: 0.75rem;
  border-color: #d2b4de;
  box-shadow: 0 4px 10px rgba(0, 0, 0, 0.04);
}
.fancy-plotcat .plotcat__button {
  border-radius: 2rem;
}
.fancy-plotcat .plotcat__compare-options {
  border-color: #d2b4de;
  background: #f5eef8;
}
</style>
```

Then apply the class to your exercise `Div`:

````markdown
::: {.plotcat .fancy-plotcat}
```{r}
plot(cars)
```
:::
````

### Result

::: {.plotcat .fancy-plotcat}
```{r}
plot(cars)
```
:::
