---
title: Reference
format:
  live-html:
    toc: true
    toc-depth: 3
---

```{=html}
<style>
.plotcat.accent-green {
  --plotcat-accent: #059669;
  --plotcat-accent-rgb: 5, 150, 105;
}

.plotcat.custom-card {
  --plotcat-bg: #fffbeb;
  --plotcat-border: #f59e0b;
  --plotcat-accent: #d97706;
  --plotcat-accent-rgb: 217, 119, 6;
}
</style>
```

Configuration options, chunk markup syntax, and CSS theme variables for the `plotcat` extension.

## Document Setup

Install Quarto Live and PlotCat in the project:

```bash
quarto add r-wasm/quarto-live
quarto add VisruthSK/PlotCat
```

Use Quarto Live's HTML format and enable PlotCat:

```yaml
format: live-html

filters:
  - plotcat
```

R pages must declare the `svglite` package under `format.live-html.webr.packages`. List every non-base R or Python package used by target or starter code. PlotCat does not infer or install runtime packages; configure them through Quarto Live.

```yaml
format:
  live-html:
    webr:
      packages:
        - svglite
        - ggplot2
    pyodide:
      packages:
        - matplotlib
```

## Exercise Markup

Write the target plot as the first chunk inside a `.plotcat` Div. Set `#| eval: false` so Quarto leaves it for the browser runtime.

````markdown
::: {.plotcat}
```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```
:::
````

Add a second chunk for starter code. Set `#| eval: false` on the starter chunk.

````markdown
::: {.plotcat}
```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```

```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars
)
```
:::
````

### Markup Rules

- Both chunks need `#| eval: false`. PlotCat executes code in the browser, not during document rendering.
- The first chunk becomes the browser-rendered target. Its source is obfuscated in the HTML. The obfuscation discourages casual inspection but is not a security boundary.
- The second chunk populates the student code editor. Omit it for a blank editor.
- Both chunks in an exercise must use the same language. Use `{r}` or `{webr}` for R, and `{python}` or `{pyodide}` for Python. PlotCat normalizes the student editor to Quarto Live's native `webr` or `pyodide` cell.
- Assign custom exercise IDs on the container div: `::: {.plotcat #scatter-exercise}`. Auto-generated IDs are `exercise-1`, `exercise-2`, etc.
- Extra classes on the `.plotcat` Div are forwarded to the widget container for scoped styling.

## Comparison Modes

SVG plots support three comparison modes. Plotly charts use side-by-side comparison; overlay and wipe are unavailable.

| Mode | Behavior |
|---|---|
| **Side by side** (default) | Target and student plots rendered in a two-column grid. Stacks to single column below 48rem. |
| **Overlay** | Student plot rendered on top of the target at 50% opacity. |
| **Wipe** | A draggable vertical slider reveals the target on the left and the student plot on the right. Keyboard: Arrow keys, Home, End. |

## Scoring & Comparison

PlotCat compares SVG plots using feature weights and Plotly figures using four weighted categories.

### SVG Weights

Set per-exercise weight overrides as Div attributes:

```markdown
::: {.plotcat id="scatter"
    svg-geometry="0.4"
    svg-text="0.4"}
```

Set document-wide defaults under `plotcat.weights.svg` in YAML frontmatter:

```yaml
plotcat:
  weights:
    svg:
      geometry: 0.5
      text: 0.25
```

| Attribute | Default | Controls |
|---|---|---|
| `svg-geometry` | 0.6 | Element positions (bag overlap of coarse geometry, coordinates rounded to ~0.1% relative precision) |
| `svg-text` | 0.15 | Text content match (bag overlap of visible `<text>` strings) |
| `svg-style` | 0.1 | Fill, stroke, opacity attributes on geometry elements |
| `svg-frame` | 0.15 | Aspect ratio from viewBox or explicit dimensions (pass/fail, not proportional) |

### Plotly figures

PlotCat compares the target and student figure JSON, rather than the rendered pixels or source code. It reads four top-level fields:

| Field | What PlotCat can compare |
|---|---|
| `data` | Trace type and order, coordinates, `customdata`, hover text, marker and line properties, and other serialized trace settings |
| `layout` | Titles, axes, ranges, legends, annotations, shapes, margins, and other serialized layout settings |
| `frames` | Names, data, and layout stored in animation frames |
| `config` | Configuration values included by the browser runtime |

PlotCat grades each property present in the target. Set a title, color, range, hover template, or similar property in the target code when it should affect the score. Plotly defaults that do not appear in the target JSON are not graded.

| Weight | Default | Covers |
|---|---:|---|
| `plotly-trace` | 0.3 | Trace count and type |
| `plotly-data` | 0.4 | All non-style trace data, including coordinates, names, hover text, and `customdata` |
| `plotly-style` | 0.2 | Trace styling, including marker, line, fill, and opacity settings |
| `plotly-layout` | 0.1 | Layout, animation frames, and Plotly configuration |

Arrays are positional. Missing, reordered, or extra traces, points, annotations, and shapes reduce the score. Extra properties inside a student object are ignored, so students may add settings that the target does not require.

PlotCat ignores generated `uid` values and keys that begin with `_`. It allows small floating-point differences. In text properties, `<b>` and `<strong>` are equivalent, as are `<i>` and `<em>` and common forms of `<br>`.

Functions, callbacks, and settings omitted from the serialized figure cannot be compared. Plotly exercises use side-by-side mode because overlay and wipe are unavailable for Plotly charts.

## Dimensions

Override plot dimensions per exercise. Default is 7 inches wide by 5 inches tall for both R and Python.

```markdown
::: {.plotcat id="tall-plot" plotcat-width="4" plotcat-height="6"}
```

Set document-wide defaults under `plotcat.dimensions` in YAML frontmatter.

```yaml
plotcat:
  dimensions:
    width: 8
    height: 6
```

## Heading

Override the heading text per exercise. Defaults to "Recreate this plot".

```markdown
::: {.plotcat id="labeled" plotcat-heading="Reproduce the graphic"}
```

Set the heading in YAML frontmatter for a document-wide default.

```yaml
plotcat:
  heading: "Reproduce the graphic"
```

## CSS Variables

PlotCat exposes CSS custom properties on `.plotcat` containers. Override these variables globally in your stylesheet or locally on specific containers.

| Variable | Default | Description |
|---|---|---|
| `--plotcat-border` | `color-mix(in srgb, var(--bs-body-color, #212529) 20%, var(--bs-body-bg, #ffffff))` | Border color for the exercise card and internal dividers |
| `--plotcat-accent` | `var(--bs-primary, #0d6efd)` | Brand color for buttons, focus highlights, score badges, active mode toggles, and the wipe handle |
| `--plotcat-accent-rgb` | `var(--bs-primary-rgb, 13, 110, 253)` | Space-separated RGB values used for focus ring and badge background opacity channels |
| `--plotcat-bg` | `var(--bs-body-bg, #ffffff)` | Card and plot container background |
| `--plotcat-nested-bg` | `color-mix(in srgb, var(--bs-body-color, #212529) 6%, var(--bs-body-bg, #ffffff))` | Surface color for the code editor and comparison control bar |
| `--plotcat-shadow-color` | `rgb(var(--bs-emphasis-color-rgb, 0, 0, 0))` | Tint for card drop shadows and handle depth |

### Bootstrap Variable Integration

PlotCat inherits global colors from Quarto Bootstrap themes.

- `--bs-body-color`: Primary text color for labels, status messages, and editor code.
- `--bs-body-bg`: Baseline for card and plot surfaces.
- `--bs-primary`: Mapped to `--plotcat-accent`.
- `--bs-primary-rgb`: Mapped to `--plotcat-accent-rgb`.
- `--bs-font-monospace`: Font stack for the code editor.
- `--bs-body-font-family`: Font stack for headers, labels, and buttons.
- `--bs-secondary-color`: Color for muted comparison labels and idle status text.
- `--bs-danger`: Color for execution error messages.

## Custom Styling Examples

### Theme Accent Override

```css
.plotcat.accent-green {
  --plotcat-accent: #059669;
  --plotcat-accent-rgb: 5, 150, 105;
}
```

::: {.plotcat .accent-green}
```{r}
#| eval: false
tinyplot::tinyplot(dist ~ speed, data = cars, main = "Stopping distance by speed", xlab = "Speed", ylab = "Stopping distance")
```

```{r}
#| eval: false
tinyplot::tinyplot(dist ~ speed, data = cars)
```
:::

### Scoped Container Styling

#### Custom card

```css
.plotcat.custom-card {
  --plotcat-bg: #fffbeb;
  --plotcat-border: #f59e0b;
  --plotcat-accent: #d97706;
  --plotcat-accent-rgb: 217, 119, 6;
}
```

::: {.plotcat .custom-card}
```{r}
#| eval: false
library(lattice)
xyplot(mpg ~ wt, data = mtcars, main = "MPG vs Weight", xlab = "Weight", ylab = "MPG")
```

```{r}
#| eval: false
library(lattice)
xyplot(mpg ~ wt, data = mtcars)
```
:::
