---
title: Reference
format:
  live-html:
    toc: true
    toc-depth: 3
---

```{=html}
<style>
.plotcat.accent-green {
  --plotcat-accent: #059669;
  --plotcat-accent-rgb: 5, 150, 105;
}

.plotcat.custom-card {
  --plotcat-bg: #fffbeb;
  --plotcat-border: #f59e0b;
  --plotcat-accent: #d97706;
  --plotcat-accent-rgb: 217, 119, 6;
}
</style>
```

Configuration options, chunk markup syntax, and CSS theme variables for the `plotcat` extension.

## Document Setup

Install Quarto Live and PlotCat in the project:

```bash
quarto add r-wasm/quarto-live
quarto add VisruthSK/PlotCat
```

Use Quarto Live's HTML format and enable PlotCat:

```yaml
format: live-html

filters:
  - plotcat
```

R pages must declare the `svglite` package under `format.live-html.webr.packages`. Every non-base R or Python package used by target or starter code must be listed too. PlotCat does not infer or install runtime packages — package configuration belongs to Quarto Live.

```yaml
format:
  live-html:
    webr:
      packages:
        - svglite
        - ggplot2
    pyodide:
      packages:
        - matplotlib
```

## Exercise Markup

Write the target plot as the first chunk inside a `.plotcat` Div. Set `#| eval: false` so Quarto leaves it for the browser runtime.

````markdown
::: {.plotcat}
```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```
:::
````

Add a second chunk for starter code. Also set `#| eval: false` on the starter chunk.

````markdown
::: {.plotcat}
```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars,
  main = "Stopping distance by speed",
  xlab = "Speed",
  ylab = "Stopping distance"
)
```

```{{r}}
#| eval: false
tinyplot::tinyplot(
  dist ~ speed,
  data = cars
)
```
:::
````

### Markup Rules

- Both chunks need `#| eval: false`. PlotCat executes code in the browser, not during document rendering.
- The first chunk becomes the browser-rendered target. Its source is obfuscated in the HTML — enough to prevent casual inspection, not a security boundary.
- The second chunk populates the student code editor. Omit it for a blank editor.
- Both chunks in an exercise must use the same language. Use `{r}` or `{webr}` for R, and `{python}` or `{pyodide}` for Python. PlotCat normalizes the student editor to Quarto Live's native `webr` or `pyodide` cell.
- Assign custom exercise IDs on the container div: `::: {.plotcat #scatter-exercise}`. Auto-generated IDs are `exercise-1`, `exercise-2`, etc.
- Extra classes on the `.plotcat` Div are forwarded to the widget container for scoped styling.

## Comparison Modes

Three comparison modes are available for SVG plots. Plotly charts are limited to side-by-side only — overlay and wipe are disabled and hidden behind a tooltip.

| Mode | Behavior |
|---|---|
| **Side by side** (default) | Target and student plots rendered in a two-column grid. Stacks to single column below 48rem. |
| **Overlay** | Student plot rendered on top of the target at 50% opacity. |
| **Wipe** | A draggable vertical slider reveals the target on the left and the student plot on the right. Keyboard: Arrow keys, Home, End. |

## Scoring & Comparison

PlotCat compares SVG plots using feature weights and Plotly figures using an explicit target contract.

### SVG Weights

Set per-exercise weight overrides as Div attributes:

```markdown
::: {.plotcat id="scatter"
    svg-geometry="0.4"
    svg-text="0.4"}
```

Set document-wide defaults under `plotcat.weights.svg` in YAML frontmatter:

```yaml
plotcat:
  weights:
    svg:
      geometry: 0.5
      text: 0.25
```

| Attribute | Default | Controls |
|---|---|---|
| `svg-geometry` | 0.6 | Element positions (bag overlap of coarse geometry, coordinates rounded to ~0.1% relative precision) |
| `svg-text` | 0.15 | Text content match (bag overlap of visible `<text>` strings) |
| `svg-style` | 0.1 | Fill, stroke, opacity attributes on geometry elements |
| `svg-frame` | 0.15 | Aspect ratio from viewBox or explicit dimensions (pass/fail, not proportional) |

### Plotly Exercises

PlotCat evaluates Plotly charts by comparing serialized figure JSON (`data`, `layout`, `frames`, `config`) directly against the target plot.

- **Explicit Property Contract**: Every property explicitly defined in the target figure must be present in the student figure. Unset target properties are not graded against implicit Plotly defaults. Explicitly define titles, axis labels, colors, ranges, and hover templates in the target code if students should be scored on them. It is not suggested to define them and set them to their default values.
- **Trace and Array Order**: Traces, data points, annotations, and shapes are matched by position. Extra student array elements reduce the score.
- **HTML Text Equivalents**: Basic markup tags in text properties (`<b>` / `<strong>`, `<i>` / `<em>`, `<br>` / `<br/>`) are treated as identical.
- **Comparison Modes**: Plotly charts render in side-by-side mode. Overlay and wipe modes are disabled for Plotly figures.

## Dimensions

Override plot dimensions per exercise. Default is 7 inches wide by 5 inches tall for both R and Python.

```markdown
::: {.plotcat id="tall-plot" plotcat-width="4" plotcat-height="6"}
```

Set document-wide defaults under `plotcat.dimensions` in YAML frontmatter.

```yaml
plotcat:
  dimensions:
    width: 8
    height: 6
```

## Heading

Override the heading text per exercise. Defaults to "Recreate this plot".

```markdown
::: {.plotcat id="labeled" plotcat-heading="Reproduce the graphic"}
```

Set the heading in YAML frontmatter for a document-wide default.

```yaml
plotcat:
  heading: "Reproduce the graphic"
```

## CSS Variables

PlotCat exposes CSS custom properties on `.plotcat` containers. Override these variables globally in your stylesheet or locally on specific containers.

| Variable | Default | Description |
|---|---|---|
| `--plotcat-border` | `color-mix(in srgb, var(--bs-body-color, #212529) 20%, var(--bs-body-bg, #ffffff))` | Border color for the exercise card and internal dividers |
| `--plotcat-accent` | `var(--bs-primary, #0d6efd)` | Brand color for buttons, focus highlights, score badges, active mode toggles, and the wipe handle |
| `--plotcat-accent-rgb` | `var(--bs-primary-rgb, 13, 110, 253)` | Space-separated RGB values used for focus ring and badge background opacity channels |
| `--plotcat-bg` | `var(--bs-body-bg, #ffffff)` | Card and plot container background |
| `--plotcat-nested-bg` | `color-mix(in srgb, var(--bs-body-color, #212529) 6%, var(--bs-body-bg, #ffffff))` | Surface color for the code editor and comparison control bar |
| `--plotcat-shadow-color` | `rgb(var(--bs-emphasis-color-rgb, 0, 0, 0))` | Tint for card drop shadows and handle depth |

### Bootstrap Variable Integration

PlotCat inherits global colors from Quarto Bootstrap themes automatically.

- `--bs-body-color`: Primary text color for labels, status messages, and editor code.
- `--bs-body-bg`: Baseline for card and plot surfaces.
- `--bs-primary`: Mapped to `--plotcat-accent`.
- `--bs-primary-rgb`: Mapped to `--plotcat-accent-rgb`.
- `--bs-font-monospace`: Font stack for the code editor.
- `--bs-body-font-family`: Font stack for headers, labels, and buttons.
- `--bs-secondary-color`: Color for muted comparison labels and idle status text.
- `--bs-danger`: Color for execution error messages.

## Custom Styling Examples

### Theme Accent Override

```css
.plotcat.accent-green {
  --plotcat-accent: #059669;
  --plotcat-accent-rgb: 5, 150, 105;
}
```

::: {.plotcat .accent-green}
```{r}
#| eval: false
tinyplot::tinyplot(dist ~ speed, data = cars, main = "Stopping distance by speed", xlab = "Speed", ylab = "Stopping distance")
```

```{r}
#| eval: false
tinyplot::tinyplot(dist ~ speed, data = cars)
```
:::

### Scoped Container Styling

#### Custom card

```css
.plotcat.custom-card {
  --plotcat-bg: #fffbeb;
  --plotcat-border: #f59e0b;
  --plotcat-accent: #d97706;
  --plotcat-accent-rgb: 217, 119, 6;
}
```

::: {.plotcat .custom-card}
```{r}
#| eval: false
library(lattice)
xyplot(mpg ~ wt, data = mtcars, main = "MPG vs Weight", xlab = "Weight", ylab = "MPG")
```

```{r}
#| eval: false
library(lattice)
xyplot(mpg ~ wt, data = mtcars)
```
:::

